Sequential algorithms to split and merge ultra-high resolution 3D images

### 2017-08-14 paper.tex

\author{Final ordering
  to be discussed: Yongping Gao/Val\'erie Hayot-Sasson,
  Yuhong Yan, Tristan Glatard\\
  Department of Computer Science and Software Engineering\\ Concordia University, Montreal, Quebec, Canada\\
  {first.last}@concordia.ca
}

@@

%% \subsection{Lossless compression}

@@

% Clustered reads
Clustered reads provide important speed ups compared to naive blocks,
both on HDD and on SSD. On HDD, they are up to 9.7 times faster than
naive blocks at 16~GB and 6.8 times faster on average. On SSD, they
are up to 5.1 times faster than naive blocks at 16~GB, and 3.1 times
faster on average.  Surprisingly, they perform substantially faster
than naive blocks even at 3~GB, while in Case 1. This may be explained
by the fact that the seeks required to write incomplete block rows to
the reconstructed image are shorter than the ones for naive blocks.

% Multiple reads
Multiple reads are even faster than Clustered reads on this dataset. On
HDD, they are 8.4 times faster than naive blocks on average (5.3 times
on SSD). At 16~GB, they are 9.0 times faster than naive blocks on
average (5.8 times faster on SSD).

@@

@@

% File formats
In terms of file formats, our results demonstrate that simple imaging
formats may be split and merged without performance loss compared to
more complex formats that try to preserve spatial locality on disk,
for instance MINC 2.0 or the format based on space-filling curves
mentioned in~\cite{burns2013open}. This is of major interest in the
current open-science context since simpler formats favor data-sharing
and interoperability.  Moreover, our algorithms could potentially be
adapted to any split geometry, even though we demonstrated them on
slices and blocks only, while file formats inevitably assume a
particular geometry. For instance, the format in~\cite{burns2013open}
is not designed to naively split slices. However, we aimed to extract
all the blocks, whereas, ~\cite{burns2013open} aimed to extract a
single block from a large image. Furthermore, we have not considered
on-the-fly data compression yet, which is a unique feature of MINC
2.0.

>>>>

\author{Val\'erie Hayot-Sasson$^*$, Yongping Gao$^*$, Yuhong Yan, Tristan Glatard\\
  Department of Computer Science and Software Engineering\\ Concordia University, Montreal, Quebec, Canada\\
  {first.last}@concordia.ca\\
  $^*$ These authors have contributed equally

}

@@

\subsection{Lossless compression}

We implemented lossless compression for all algorithms using Python's
gzip library. Compression is done on the fly, that is, while the data
is being read or written. On-the-fly compression of large datasets is
a challenge when extensive seeking is involved as the gzip library has
to decompress all the data until the seek position to read from it,
making access time a linear function of the seek position.

@@

% Clustered reads
Clustered reads provide important speed ups compared to naive blocks,
both on HDD and on SSD. They are 6.8 times faster than naive blocks on
HDD and 5.1 times on SSD (average accross all repetitions, all memory
values). Surprisingly, they perform substantially faster than naive
blocks even at 3~GB, while in Case 1. This may be explained by the
fact that the seeks required to write incomplete block rows to the
reconstructed image are shorter than the ones for naive blocks.

% Multiple reads
Multiple reads are even faster than Clustered reads on this
dataset. They are 8.4 times faster than naive blocks on HDD and 5.3
times on SSD (average accross all repetitions, all memory values).

@@

\subsection{Compression}

Naive blocks, naive slices, cluster reads and multiple reads at 3, 6,
9, 12, 16GB, on SSD, with on-the-fly compression (graph will also show
without compression (previous results)) for comparison.

Expected result: Compression speeds up Multiple reads. It speeds up
Clustered reads to a smaller extent since Cluster reads still have to
seek in the reconstructed image. \todo{finish that up when the data is
  available}

@@

% File formats
In terms of file formats, our results demonstrate that simple imaging
formats may be split and merged without performance loss compared to
more complex formats that try to preserve spatial locality on disk,
for instance MINC 2.0 or the format based on space-filling curves
mentioned in~\cite{burns2013open}. This is of major interest in the
current open-science context since simpler formats favor data-sharing
and interoperability.  Moreover, our algorithms could potentially be
adapted to any split geometry, even though we demonstrated them on
slices and blocks only, while file formats inevitably assume a
particular geometry. For instance, the format in~\cite{burns2013open}
is not designed to naively split slices. However, we aimed to extract
all the blocks, whereas, ~\cite{burns2013open} aimed to extract a
single block from a large image. MINC 2.0 might also help with
on-the-fly compression.

% Compression
Due to the minimal number of seeks in the reconstructed image,
Multiple reads greatly benefit from on-the-fly compression. This is
not the case of Clustered reads, where compression has an important
overhead due to residual seeking in the reconstructed
image. Techniques such as z-indexing have been used to address this
limitation of the NiFTI format~\cite{rajna2015speeding} and could be
employed here too.
